---
title: "Data Wrangling"
author: "Kasper Valtonen"
---


Challenge:
```{r}
# Setting up libraries
library(vroom)
library(tidyverse)
library(readxl)
library(data.table)
library(dplyr)

```

Let's import the reduced data to R with vroom
```{r}
#Importing data
col_types <- list(
  patent_id = col_character(),
  id = col_character(),
  assignee_id = col_character(),
  type = col_integer(),
  date = col_date("%Y-%m-%d"),
  num_claims = col_integer(),
  organization = col_character(),
  mainclass_id = col_character(),
  sequence = col_integer()
)

patent_tbl <- vroom(
            file       = "C:\\Users\\35844\\ML\\ss23-bdsb-TurvaKassu\\Patent_data_reduced\\patent.tsv", 
            delim      = "\t", 
            col_types  = col_types,
            na         = c("", "NA", "NULL")
        )

patent_assignee_tbl <- vroom(
            file = "C:\\Users\\35844\\ML\\ss23-bdsb-TurvaKassu\\Patent_data_reduced\\patent_assignee.tsv",
            delim      = "\t", 
            col_types  = col_types,
            na         = c("", "NA", "NULL")
            )

assignee_tbl <- vroom(
            file = "C:\\Users\\35844\\ML\\ss23-bdsb-TurvaKassu\\Patent_data_reduced\\assignee.tsv",
            delim      = "\t", 
            col_types  = col_types,
            na         = c("", "NA", "NULL")
            )

uspc_tbl <- vroom(
            file = "C:\\Users\\35844\\ML\\ss23-bdsb-TurvaKassu\\Patent_data_reduced\\uspc.tsv",
            delim      = "\t", 
            col_types  = col_types,
            na         = c("", "NA", "NULL")
            )

# Combine tibbles
complete_patent_tbl <- left_join(
  patent_tbl, patent_assignee_tbl, by = c("id" = "patent_id")) %>%
  left_join(assignee_tbl, by = c("assignee_id" = "id")) %>%
  left_join(uspc_tbl, by = c("id" = "patent_id"))
```

Transform tibble into data.table and excluding NA values for organization
```{r}
patent_dt <- as.data.table(complete_patent_tbl)[!is.na(organization)][!is.na(mainclass_id)]
```

Question 1: Which organizations posses most number of patents?
```{r}
number_of_patents <- patent_dt[, .N, by=organization] %>%
  .[order(-N)] %>%
  .[1:10]
print(number_of_patents)
  
```

Question 2: Companies with most patents in August 2014.

First we need to add date column to previous dt manipulation
```{r}
number_of_patents_august <- patent_dt[year(date) == 2014 & month(date) == 8, .N, by=organization] %>%
  .[order(-N)] %>%
  .[1:10]
print(number_of_patents)

```

Question 3: What is the most innovative tech industry?

Solve the main class of most number of patents
```{r}
# Get the top10 orgs with most patents
top10_orgs <- patent_dt[, .N, by=organization] %>%
  .[order(-N)] %>%
  .[1:10] %>%
  select(organization)

# Only include rows with orgs from top10_orgs and group by mainclass_id
top5_mainclasses <- patent_dt[organization %in% top10_orgs$organization, 
                              .N, by=mainclass_id] %>%
                    .[order(-N)] %>%
                    .[1:5]

print(top5_mainclasses)

```
